A Model to Predict Survival in Patients with End-Stage Liver Disease

### 2021-10-15 analysis/article.Rmd
fix: gtsummary compact theme
feat: add ml method part

```{r knitr_setup, include=FALSE}
knitr::write_bib(
    c("targets", "workflowr"),
    file = rprojroot::find_rstudio_root_file(
        "analysis", "bibliography", "rpackages", "article.bib"
    )
)
```

```{r article_setup, include=FALSE}
library("targets")
library("ameld")

tar_load(bmrk_results)

@@

```{r ml, include = FALSE}
n_lrns <- nrow(bmrk_results$learners)
#n_cv <- c(inner.folds
#    outer = unlist(
#        bmrk_results$resamplings$resampling[[1]]$param_set$get_values()
#    )
#)
```

@@

```{r tbl1, results = "asis", echo = FALSE, message = FALSE}
tbl1
```

@@

>>>>

```{r knitr_setup, include=FALSE}
knitr::write_bib(
    c(
        "base",
        "targets", "workflowr",
        "mlr3", "mlr3proba", "zlog",
        "survival", "glmnet", "penalized", "ranger", "randomForestSRC",
        "xgboost", "survivalsvm", "survivalmodels"
    ),
    file = rprojroot::find_rstudio_root_file(
        "analysis", "bibliography", "rpackages", "article.bib"
    )
)
```

```{r article_setup, include=FALSE}
library("targets")
library("ggplot2")
library("mlr3viz")
library("viridis")
library("ameld")

tar_load(bmrk_results)
tar_load(crossval)

@@

```{r ml, include = FALSE}
n_lrns <- sum(!grepl("^scale..*", bmrk_results$learners$learner_id))
#n_cv <- c(inner.folds
#    outer = unlist(
#        bmrk_results$resamplings$resampling[[1]]$param_set$get_values()
#    )
#)
```

All statistical and machine learning analyses were performed using R version
`r paste0(sessionInfo()$R.version[c("major", "minor")], collapse = ".")`
[@R-base].
Prior to the analyses all laboratory values were zlog transformed
[@hoffmann2017; @R-zlog].
We compared `r n_lrns` different statistical and machine learning
algorithms using the mlr3 framework
[@mlr3; @R-mlr3; @mlr3proba2021; @R-mlr3proba].
The used algorithms are designed for analysis of survival data:
Cox proportional hazards regression model
[@cox1972; @survival-book; @R-survival],
penalized regressions,
namely two different implementations of the
lasso (least absolute shrinkage and selection operator) regression
[@tibshirani1997; @glmnet2011; @R-glmnet; @penalized2010; @R-penalized],
tow different implementations of the
ridge regression [@glmnet2011; @R-glmnet; @penalized2010; @R-penalized], and
the elastic net regression [@glmnet2011; @R-glmnet],
two different random forest implementations
[@ranger2017; @R-ranger; @randomForestSRC2008; @R-randomForestSRC],
extreme gradient boosting [@chen2016; @R-xgboost],
two different support vector machines [@vanbelle2011; @R-survivalsvm]
and a neural network [@katzman2018; @R-survivalmodels].

To choose the best model we used nested resampling to avoid feature- and
model-selection-bias.
Therefore, we applied a `r crossval$inner$param_set$values$folds`-fold inner
and a `r crossval$outer$param_set$values$folds`-fold outer cross-validation
that was repeated `r crossval$outer$param_set$values$repeats` times.
Subsequently we selected the model with the
highest median concordance index [@harrell1982].

TODO: selected final model, analysis, feature selection

@@

```{r tbl1, results = "asis", echo = FALSE, message = FALSE}
library("gtsummary")
theme_gtsummary_journal(journal = "jama")
theme_gtsummary_compact()
tbl1
```

@@

```{r benchmark-plot, fig.cap = "Benchmark results of machine learning algorithms."}
r <- bmrk_results$clone()
id <- grep("^scale", x = r$learners$learner_id, invert = TRUE, value = TRUE)
r$filter(task_id = "zlog_eldd", learner_id = id)
autoplot(r) +
    geom_boxplot(aes(fill = learner_id)) +
    geom_jitter(position = position_jitter(0.2)) +
    scale_fill_viridis(discrete = TRUE) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

TODO: mention best ml algorithm (\@ref(fig:benchmark-plot))
